---
title: "4PopSims"
author: "Jennifer Blanc"
date: "2024-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(latex2exp)
library(ggpubr)
```

```{r}
df <- fread("../plots/4PopSplit/B/q_extend.txt")
df <- df [,3:ncol(df)]
```

```{r}

dfPlot <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`nc-q`),lowerci_mean = avg_q - (1.96 * (sd(`nc-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`nc-q`)  /  sqrt(n()))), fp_strat = sum(`nc-p` < 0.05)/ n(), lowerci_fp = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci_fp = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), q_std = avg_q / sd(`nc-q`))
```


```{r}
pl <- dfPlot %>% filter(config =="C1", gwas_size == 10000) %>% ggplot(aes(x = env, y = fp_strat, color = type)) + geom_point(size = 2) + facet_grid(vars(L), vars(test_size)) + geom_errorbar(aes(ymin = lowerci_fp, ymax = upperci_fp), width = 0.005) + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + ylim(-0.1,1) + labs(colour="Covariate")+ geom_hline(yintercept = 0.05, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("False Postive Rate") + ggtitle("Test Panel Size")
pl

#ggsave("~/Desktop/fp.png", pl)
```

```{r}
pl <- dfPlot %>% filter(config =="C1", gwas_size == 10000) %>% ggplot(aes(x = env, y = avg_q, color = type)) + geom_point(size = 2) + facet_grid(vars(L), vars(test_size)) + geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.1) + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + labs(colour="Covariate")+ geom_hline(yintercept = 0, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("Bias q") + ggtitle("Test Panel Size")
pl

#ggsave("~/Desktop/bias.png", pl)
```

```{r}
dfPlot %>% filter(config =="C1", gwas_size == 10000) %>% ggplot(aes(x = env, y = q_std, color = type)) + geom_point(size = 2) + facet_grid(vars(L), vars(test_size)) + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + labs(colour="Covariate")+ geom_hline(yintercept = 0, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("Bias q") + ggtitle("Test Panel Size")

```



```{r}
dfE <- fread("../plots/4PopSplit/B/error.txt")
dfE <- dfE[,3:ncol(dfE)]

dfSumE <- dfE %>% group_by(config, gwas_size, test_size, L, type) %>% summarize(count = n(), avg_error = mean(error),lowerci = avg_error - (1.96 * (sd(error)  /  sqrt(n()))), upperci = avg_error + (1.96 * (sd(error)  /  sqrt(n()))))

pl <- dfSumE %>% filter(config =="C1") %>% ggplot(aes(x = L, y = avg_error, color = type)) + geom_point(size = 2) + facet_grid(vars(gwas_size), vars(test_size)) + scale_x_log10() + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width = 0.25) + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5), plot.tag.position = c(1.03, .55),plot.tag = element_text(angle = -90) ,legend.position = "bottom", plot.margin = margin(0,1.5,0,0, "cm")) + ylab("Avg. Error") + ggtitle("Test Panel Size") + labs(tag = "GWAS Panel size", colour="Covariate") + geom_hline(yintercept = 0, color = "red") + ylim(c(0,1))

pl
#ggsave("~/Desktop/error.png", pl, height = 7, width = 10)
```

## Figure 1 

```{r}
df <- fread("../plots/4PopSplit/A/q.txt")
df <- df [,3:ncol(df)]

dfPlot <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`nc-q`),lowerci_mean = avg_q - (1.96 * (sd(`nc-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`nc-q`)  /  sqrt(n()))), fp_strat = sum(`nc-p` < 0.05)/ n(), lowerci_fp = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci_fp = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), q_std = avg_q / sd(`nc-q`))

dfPlot %>% filter(config =="C1", gwas_size == 10000) %>% ggplot(aes(x = env, y = q_std, color = type)) + geom_point(size = 2) + facet_grid(vars(L), vars(test_size)) + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + labs(colour="Covariate")+ geom_hline(yintercept = 0, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("Bias q") + ggtitle("Test Panel Size")

```




## Figure 2

### Function to make Error vs SNP/Individual ratio plot

```{r}
summarize_error <- function(dfError) {
  
   # Compute SNP to individual ratio
  dfError$gamma <- dfError$L / 1000

  # Compute average accross replicates
  dfError_Plot <- dfError %>% group_by(config, test_size, gwas_size, L, type, gamma) %>% summarise(num = n(), avg_error = mean(error), lowerci_mean = avg_error - (1.96 * (sd(error)  /  sqrt(n()))), upperci_mean = avg_error + (1.96 * (sd(error)  /  sqrt(n()))))
  
  return(dfError_Plot)
}


plot_error <- function(dfError_Plot) {

  # Plot 
  pl <-ggplot(data = dfError_Plot, aes(x = gamma, y = avg_error, color = type)) + geom_point(size=2) +
  geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.1) + scale_x_log10() +
  theme_bw(base_size = 14) + theme(legend.position = "bottom", panel.border = element_rect(colour = "black", fill=NA, size=1), legend.text = element_text(size = 10), panel.grid.minor = element_line(linetype = 'dashed')) + xlab(TeX("$\\frac{L}{M}$")) + ylab("Error") + ylim(0,1) + scale_color_manual(values = c("navy", "red4"), labels = c(TeX("$\\hat{F}_{Gr}$"),"PC 1"), name = "")
  
  return(pl)
}

```

## Function to make Bias vs SNP/Individual ratio plot

```{r}
# Summarize q data
compute_summary <- function(df) {
  
  # Compute average bias across replicates 
  dfPlot <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`nc-q`),lowerci_mean = avg_q - (1.96 * (sd(`nc-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`nc-q`)  /  sqrt(n()))), fp_strat = sum(`nc-p` < 0.05)/ n(), lowerci_fp = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci_fp = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), q_std = avg_q / sd(`nc-q`), avg_q_stan = mean(`stan-nc`))

  # Compute SNP to individual ratio
  dfPlot$gamma <- dfPlot$L / 1000
  
  return(dfPlot)
}

# 1) Computing the expected bias per replicate than averaging 
opt1 <- function(dfPlot, dfError) {
  
  # Fix labeling
  dfError$type <- gsub("PC1", "1", dfError$type)
  tmp <- df %>% filter(type == "1" | type == "FGr")

  # Join dataframes
  dfComb <- inner_join(tmp, dfError)
  dfComb$exp_bias <- dfComb$`nc-q` * dfComb$error

  # Compute expected bias 
  dfExp <- dfComb %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type, gamma) %>% summarise(mean_exp_bias = mean(exp_bias), count = n())
  
  return(dfExp)
}

# 2) Computing the expected bias once after averaging 
opt2 <- function(dfPlot, dfError_Plot) {

  # Compute Expected bias based on error
  noCor <- dfPlot %>% filter(type == "NoCorrection") 
  noCor <- noCor[, c("avg_q", "gamma")]

  # Compute expected bias for FGr
  FGr <- dfError_Plot %>% filter(type == "FGr")
  FGr <- FGr[, c("avg_error", "gamma")]
  FGr$mean_exp_bias <- FGr$avg_error * noCor$avg_q
  FGr$type <- "FGr"

  # Compute expected bias for PC 1
  PC1 <- dfError_Plot %>% filter(type == "PC1")
  PC1 <- PC1[, c("avg_error", "gamma")]
  PC1$mean_exp_bias <- PC1$avg_error * noCor$avg_q
  PC1$type <- "1" 

  # Combine Results 
  dfExp <- rbind(PC1, FGr)  
  
  return(dfExp)
}

plot_bias <- function(dfPlot, dfExp) {
  
  # Fix order
  dfPlot$type <- factor(dfPlot$type, levels = c("FGr", "1", "ID" , "NoCorrection"))
  dfExp$type <- factor(dfExp$type, levels = c("FGr", "1"))

  # Plot 
  pl <- ggplot(data = dfPlot, aes(x = gamma, y = avg_q)) + geom_point(aes(color = type, alpha = type), size=2) + scale_x_log10() + 
  geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean, color = type, alpha = type), width = 0.1) + 
  xlab(TeX("$\\frac{L}{M}$")) + ylab("Bias") +  
  theme_bw(base_size = 14) + theme(legend.position = "bottom", panel.border = element_rect(colour = "black", fill=NA, size=1), legend.text = element_text(size = 10), panel.grid.minor = element_line(linetype = 'dashed')) +
  scale_color_manual(values = c("navy", "red4", "darkgreen", "goldenrod2"), labels = c(TeX("$\\hat{F}_{Gr}$"),"PC 1", TeX("$\\tilde{F}_{Gr}$"), "No Correction"), name = "Bias") + scale_alpha_manual(values = c(1,1,0.65,0.65), guide = "none")
#  + geom_point(data = dfExp, aes(x = gamma, y = mean_exp_bias, shape = type), color = "red", size = 2) +scale_shape_manual(values = c(8, 4), labels = c(TeX("$\\hat{F}_{Gr}$"), "PC 1"), name = "Expected Bias")
  
  return(pl)
}

```

### Results

**Deep Split**

First Split: 300  
Second Split: 250  


Error vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results
dfError <- fread("../plots/4PopSplit/C/error.txt")[,3:9]
dfError$gamma <- dfError$L / 1000

# Make plot
dfError_Plot <- summarize_error(dfError)
pl <- plot_error(dfError_Plot)
pl

#ggsave("~/Desktop/deep_split_error.png", pl, width = 4, height = 3)
```

Bias vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results 
df <- fread("../plots/4PopSplit/C/q_env.txt")
df <- df [,3:ncol(df)]

# Summarize data 
dfPlot <- compute_summary(df)
#dfExp <- opt1(dfPlot, dfError)
dfExp <- opt2(dfPlot, dfError_Plot)

# Plot Data
pl <- plot_bias(dfPlot, dfExp)
pl

#ggsave("~/Desktop/deep_split_q.png", pl, width = 3.95, height = 2.95)
```

**Middle Split**

First Split: 300  
Second Split: 100  

Error vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results
dfError <- fread("../plots/4PopSplit/D/error.txt")[,3:9]
dfError$gamma <- dfError$L / 1000

# Make plot
dfError_Plot <- summarize_error(dfError)
pl <- plot_error(dfError_Plot)
pl

#ggsave("~/Desktop/middle_split_error.png", pl, width = 4, height = 3)
```

Bias vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results 
df <- fread("../plots/4PopSplit/D/q.txt")
df <- df [,3:ncol(df)]

# Summarize data 
dfPlot <- compute_summary(df)
#dfExp <- opt1(dfPlot, dfError)
dfExp <- opt2(dfPlot, dfError_Plot)

# Plot Data
pl <-plot_bias(dfPlot, dfExp)
pl

#ggsave("~/Desktop/middle_split_q.png", pl, width = 3.95, height = 2.95)
```

**No Split**

First Split: 300  
Second Split: 0  

Error vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results
dfError <- fread("../plots/4PopSplit/E/error.txt")[,3:9]
dfError$gamma <- dfError$L / 1000

# Make plot
dfError_Plot <- summarize_error(dfError)
pl <- plot_error(dfError_Plot)
pl

#ggsave("~/Desktop/np_split_error.png", pl, width = 4, height = 3)
```

Bias vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results 
df <- fread("../plots/4PopSplit/E/q.txt")
df <- df [,3:ncol(df)]

# Summarize data 
dfPlot <- compute_summary(df)
#dfExp <- opt1(dfPlot, dfError)
dfExp <- opt2(dfPlot, dfError_Plot)

# Plot Data
pl <-plot_bias(dfPlot, dfExp)
pl

#ggsave("~/Desktop/no_split_q.png", pl, width = 3.95, height = 2.95)
#ggsave("~/Desktop/legend.png", pl, width = 5, height = 2.95)
```























