---
title: "4PopSims"
author: "Jennifer Blanc"
date: "2024-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(latex2exp)
library(ggpubr)
library(ggh4x)
```


## Figure 1 

```{r}
df <- fread("../plots/4PopSplit/A/q_Fig1_C1.txt")
df <- df [,3:ncol(df)]

# Summarize data
dfPlot <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`nc-q`),lowerci_mean = avg_q - (1.96 * (sd(`nc-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`nc-q`)  /  sqrt(n()))), fp_strat = sum(`nc-p` < 0.05)/ n(), lowerci_fp = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci_fp = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), q_std = avg_q / sd(`nc-q`))

# Plot Data C1 no correction
pl1 <- dfPlot %>% filter(config =="C1" & type == "NoCorrection") %>% ggplot(aes(x = env, y = avg_q)) + geom_point(size = 3, color = "goldenrod2") + theme_classic(base_size = 16)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + labs(colour="Covariate")+ geom_hline(yintercept = 0, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("Bias") + geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.005, color = "goldenrod2") + ylim(-0.02, 0.22)
pl1
#ggsave("~/Desktop/C1-NC.png", pl1, height = 3.5, width = 4.3)


# Plot Data C2 \tidle{F}Gr
pl2 <- dfPlot %>% filter(config =="C1" & type == "ID") %>% ggplot(aes(x = env, y = avg_q)) + geom_point(size = 3, color = "darkgreen") + theme_classic(base_size = 16)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + labs(colour="Covariate")+ geom_hline(yintercept = 0, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("Bias") + geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.005, color = "darkgreen") + ylim(-0.02, 0.22)
pl2
#ggsave("~/Desktop/C1-ID.png", pl2, height = 3.5, width = 4.3)
```

```{r}
df <- fread("../plots/4PopSplit/A/q_Fig1_C2.txt")
df <- df [,3:ncol(df)]

# Summarize data
dfPlot <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`nc-q`),lowerci_mean = avg_q - (1.96 * (sd(`nc-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`nc-q`)  /  sqrt(n()))), fp_strat = sum(`nc-p` < 0.05)/ n(), lowerci_fp = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci_fp = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), q_std = avg_q / sd(`nc-q`))

# Plot Data C1 no correction
pl1 <- dfPlot %>% filter(config =="C2" & type == "NoCorrection") %>% ggplot(aes(x = env, y = avg_q)) + geom_point(size = 3, color = "goldenrod2") + theme_classic(base_size = 16)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + labs(colour="Covariate")+ geom_hline(yintercept = 0, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("Bias") + geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.005, color = "goldenrod2") + ylim(-0.02, 0.22)
pl1
#ggsave("~/Desktop/C2-NC.png", pl1, height = 3.5, width = 4.3)


# Plot Data C2 \tidle{F}Gr
pl2 <- dfPlot %>% filter(config =="C2" & type == "ID") %>% ggplot(aes(x = env, y = avg_q)) + geom_point(size = 3, color = "darkgreen") + theme_classic(base_size = 16)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), plot.title = element_text(hjust = 0.5)) + labs(colour="Covariate")+ geom_hline(yintercept = 0, color = "red") + xlab(TeX("$\\Delta_{AB}$")) + ylab("Bias") + geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.005, color = "darkgreen") + ylim(-0.02, 0.22)
pl2
#ggsave("~/Desktop/C2-ID.png", pl2, height = 3.5, width = 4.3)
```


## Figure 2

### Function to make Error vs SNP/Individual ratio plot

```{r}
summarize_error <- function(dfError) {
  
   # Compute SNP to individual ratio
  dfError$gamma <- dfError$L / 1000

  # Compute average accross replicates
  dfError_Plot <- dfError %>% group_by(config, test_size, gwas_size, L, type, gamma) %>% summarise(num = n(), avg_error = mean(error), lowerci_mean = avg_error - (1.96 * (sd(error)  /  sqrt(n()))), upperci_mean = avg_error + (1.96 * (sd(error)  /  sqrt(n()))))
  
  return(dfError_Plot)
}


plot_error <- function(dfError_Plot) {

  # Plot 
  pl <-ggplot(data = dfError_Plot, aes(x = gamma, y = avg_error, color = type)) + geom_point(size=2) +
  geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.1) + scale_x_log10() +
  theme_bw(base_size = 14) + theme(legend.position = "bottom", panel.border = element_rect(colour = "black", fill=NA, size=1), legend.text = element_text(size = 10), panel.grid.minor = element_line(linetype = 'dashed')) + xlab(TeX("$\\frac{L}{M}$")) + ylab("Error") + ylim(0,1) + scale_color_manual(values = c("navy", "red4"), labels = c(TeX("$\\hat{F}_{Gr}$"),"PC 1"), name = "")
  
  return(pl)
}

```

### Function to make Bias vs SNP/Individual ratio plot

```{r}
# Summarize q data
compute_summary <- function(df) {
  
  # Compute average bias across replicates 
  dfPlot <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`nc-q`),lowerci_mean = avg_q - (1.96 * (sd(`nc-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`nc-q`)  /  sqrt(n()))), fp_strat = sum(`nc-p` < 0.05)/ n(), lowerci_fp = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci_fp = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), q_std = avg_q / sd(`nc-q`), avg_q_stan = mean(`stan-nc`))

  # Compute SNP to individual ratio
  dfPlot$gamma <- dfPlot$L / 1000
  
  return(dfPlot)
}

# 1) Computing the expected bias per replicate than averaging 
opt1 <- function(dfPlot, dfError) {
  
  # Fix labeling
  dfError$type <- gsub("PC1", "1", dfError$type)
  tmp <- df %>% filter(type == "1" | type == "FGr")

  # Join dataframes
  dfComb <- inner_join(tmp, dfError)
  dfComb$exp_bias <- dfComb$`nc-q` * dfComb$error

  # Compute expected bias 
  dfExp <- dfComb %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type, gamma) %>% summarise(mean_exp_bias = mean(exp_bias), count = n())
  
  return(dfExp)
}

# 2) Computing the expected bias once after averaging 
opt2 <- function(dfPlot, dfError_Plot) {

  # Compute Expected bias based on error
  noCor <- dfPlot %>% filter(type == "NoCorrection") 
  noCor <- noCor[, c("avg_q", "gamma")]

  # Compute expected bias for FGr
  FGr <- dfError_Plot %>% filter(type == "FGr")
  FGr <- FGr[, c("avg_error", "gamma")]
  FGr$mean_exp_bias <- FGr$avg_error * noCor$avg_q
  FGr$type <- "FGr"

  # Compute expected bias for PC 1
  PC1 <- dfError_Plot %>% filter(type == "PC1")
  PC1 <- PC1[, c("avg_error", "gamma")]
  PC1$mean_exp_bias <- PC1$avg_error * noCor$avg_q
  PC1$type <- "1" 

  # Combine Results 
  dfExp <- rbind(PC1, FGr)  
  
  return(dfExp)
}

plot_bias <- function(dfPlot, dfExp) {
  
  # Fix order
  dfPlot$type <- factor(dfPlot$type, levels = c("FGr", "1", "ID" , "NoCorrection"))
  dfExp$type <- factor(dfExp$type, levels = c("FGr", "1"))

  # Plot 
  pl <- ggplot(data = dfPlot, aes(x = gamma, y = avg_q)) + geom_point(aes(color = type, alpha = type), size=2) + scale_x_log10() + 
  xlab(TeX("$\\frac{L}{M}$")) + ylab("Bias") +  
  theme_bw(base_size = 14) + theme(legend.position = "bottom", panel.border = element_rect(colour = "black", fill=NA, size=1), legend.text = element_text(size = 10), panel.grid.minor = element_line(linetype = 'dashed')) +
  scale_color_manual(values = c("navy", "red4", "darkgreen", "goldenrod2"), labels = c(TeX("$\\hat{F}_{Gr}$"),"PC 1", TeX("$\\tilde{F}_{Gr}$"), "No Correction"), name = "Bias") + scale_alpha_manual(values = c(1,1,0.65,0.65), guide = "none")+ geom_errorbar(aes(ymin = lowerci_mean, ymax = upperci_mean, color = type, alpha = type), width = 0.1)
  #+ geom_point(data = dfExp, aes(x = gamma, y = mean_exp_bias, shape = type), color = "red", size = 2) +scale_shape_manual(values = c(8, 4), labels = c(TeX("$\\hat{F}_{Gr}$"), "PC 1"), name = "Expected Bias") 
  
  return(pl)
}

```

### Results

**Deep Split**

First Split: 300  
Second Split: 250  


Error vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results
dfError <- fread("../plots/4PopSplit/C/error.txt")[,3:9]
dfError$gamma <- dfError$L / 1000

# Make plot
dfError_Plot <- summarize_error(dfError)
pl <- plot_error(dfError_Plot)
pl

#ggsave("~/Desktop/deep_split_error.png", pl, width = 4, height = 3)
```

Bias vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results 
df <- fread("../plots/4PopSplit/C/q_env.txt")
df <- df [,3:ncol(df)]

# Summarize data 
dfPlot <- compute_summary(df)
#dfExp <- opt1(dfPlot, dfError)
dfExp <- opt2(dfPlot, dfError_Plot)

# Plot Data
pl <- plot_bias(dfPlot, dfExp)
pl

#ggsave("~/Desktop/deep_split_q.png", pl, width = 3.95, height = 2.95)
```

**Middle Split**

First Split: 300  
Second Split: 100  

Error vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results
dfError <- fread("../plots/4PopSplit/D/error.txt")[,3:9]
dfError$gamma <- dfError$L / 1000

# Make plot
dfError_Plot <- summarize_error(dfError)
pl <- plot_error(dfError_Plot)
pl

#ggsave("~/Desktop/middle_split_error.png", pl, width = 4, height = 3)
```

Bias vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results 
df <- fread("../plots/4PopSplit/D/q.txt")
df <- df [,3:ncol(df)]

# Summarize data 
dfPlot <- compute_summary(df)
#dfExp <- opt1(dfPlot, dfError)
dfExp <- opt2(dfPlot, dfError_Plot)

# Plot Data
pl <-plot_bias(dfPlot, dfExp)
pl

#ggsave("~/Desktop/middle_split_q.png", pl, width = 3.95, height = 2.95)
```

**No Split**

First Split: 300  
Second Split: 0  

Error vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results
dfError <- fread("../plots/4PopSplit/E/error.txt")[,3:9]
dfError$gamma <- dfError$L / 1000

# Make plot
dfError_Plot <- summarize_error(dfError)
pl <- plot_error(dfError_Plot)
pl

#ggsave("~/Desktop/np_split_error.png", pl, width = 4, height = 3)
```

Bias vs SNP/Individual Ratio 
```{r, warning=FALSE, message=FALSE}
# Read in results 
df <- fread("../plots/4PopSplit/E/q.txt")
df <- df [,3:ncol(df)]

# Summarize data 
dfPlot <- compute_summary(df)
#dfExp <- opt1(dfPlot, dfError)
dfExp <- opt2(dfPlot, dfError_Plot)

# Plot Data
pl <-plot_bias(dfPlot, dfExp)
pl

#ggsave("~/Desktop/no_split_q.png", pl, width = 3.95, height = 2.95)
#ggsave("~/Desktop/legend.png", pl, width = 5, height = 2.95)
```


## Figure 3

```{r}
# Read in results 
df <- fread("../plots/4PopSplit/A/q_causal.txt")
df <- df [,3:ncol(df)]

# Summarize data 
dfPlot_nc <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`nc-q`),lowerci_mean = avg_q - (1.96 * (sd(`nc-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`nc-q`)  /  sqrt(n()))), avg_true = mean(`true-q`))
dfPlot_nc$effects <- "Ascertained" 
dfPlot_c <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L, type) %>% summarize(count = n(), avg_q = mean(`c-q`),lowerci_mean = avg_q - (1.96 * (sd(`c-q`)  /  sqrt(n()))), upperci_mean = avg_q + (1.96 * (sd(`c-q`)  /  sqrt(n()))), avg_true = mean(`true-q`))
dfPlot_c$effects <- "Causal"
dfPlot <- rbind(dfPlot_nc, dfPlot_c)

# Get true q 
dfTrue <- df %>% group_by(config, gwas_size, test_size, h2, ts, num_causal, env, L) %>% summarize(count = n(), avg_q = mean(`true-q`), avg_true = mean(`true-q`))
dfTrue$type <- "True Effect Size" 
dfTrue_nc <- dfTrue
dfTrue_nc$effects <- "Ascertained"
dfTrue_c <- dfTrue
dfTrue_c$ effects <- "Causal"
dfTrue <- rbind(dfTrue_nc, dfTrue_c)

# Final plotting data set
dfFinal <- rbind(dfPlot, dfTrue)

# Order factors
dfFinal$type <- factor(dfFinal$type, levels = c("1", "FGr", "ID" , "NoCorrection", "True Effect Size"))

# Fix facet names 
dfFinal <- dfFinal %>% mutate(direction = case_when(env == -0.1 ~ "Opposite Direction", env == 0 ~ "No Environmental Stratification", env == 0.1 ~ "Same Direction"))
dfFinal$facet2 <- factor(dfFinal$direction, levels = c("Opposite Direction","No Environmental Stratification","Same Direction" ))
dfFinal$effects <- factor(dfFinal$effects, levels = c("Causal","Ascertained"))

# Pull out just estimated effect sizes for error bars
dfEB <- dfFinal %>% filter(type %in% c("1", "FGr", "ID" , "NoCorrection"))

# Plot data 
labs <-  c("PC 1",TeX("$\\hat{F}_{Gr}$"), TeX("$\\tilde{F}_{Gr}$"), "No Correction" ,"True Effect Size")
ggplot(data = dfFinal, aes(x = avg_true, y = avg_q, color = type, shape = type)) + geom_point(aes(size = type)) + facet_grid(vars(effects), vars(facet2), scales = "free_y") + geom_errorbar(data = dfEB, aes(ymin = lowerci_mean, ymax = upperci_mean), width = 0.01, alpha = 0.5) + theme_classic(base_size = 12) + 
  theme(legend.text.align = 0)+ geom_hline(yintercept = 0, color = "red") + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.title = element_text(size = 12)) +
  scale_colour_manual(name = "",labels = labs, values = c("red4" , "navy",  "darkgreen", "goldenrod2", "red")) + 
  scale_shape_manual(name = "",labels = labs, values = c(16, 16, 16,16, 8)) + 
  scale_size_manual(name = "",labels = labs,values = c(2, 2, 2, 2, 4))

```





















